---
title: "Assignment K-means Clustering"
author: "Jyoti Phogat"
date: "2022-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Hmisc) #Contents and Describe
library(leaps) #Variable selection
library(MASS)
library(NbClust)
```

```{r}
pharmaceutical_data <- read.csv("Pharmaceuticals.csv", header=TRUE)
pharmaceutical_data
pc<-pharmaceutical_data[,2]
row.names(pharmaceutical_data) <- pharmaceutical_data[,2]
pharmaceutical_data <- pharmaceutical_data[, -c(1,2,12,13,14)]
pharmaceutical_data.norm <- sapply(pharmaceutical_data, scale)
set.seed(42)

devAskNewPage(ask=TRUE)

nc <- NbClust(pharmaceutical_data.norm, min.nc=2, 
              max.nc=10, method="kmeans")
table(nc$Best.n[1,])

barplot(table(nc$Best.n[1,]), xlab="Number of Clusters", ylab="Number of criteria", main="Number of clusters chosen by criteria")
```

#Use the wss plot to check the number of clusters

```{r}
wssplot <- function(pharmaceutical_data.norm, nc=10, seed=42) {
  wss <- (nrow(pharmaceutical_data.norm)-1)*sum(apply(pharmaceutical_data.norm, 2, var)) 
  for (i in 2:nc) {
    set.seed(42) 
    wss[i] <- sum(kmeans(pharmaceutical_data.norm, centers=i)$withinss)
  } 
  plot(1:nc, wss, type="b", xlab="Number of clusters", ylab="Within groups sum of squares")
}
wssplot(pharmaceutical_data.norm,nc=10)
```

# Perform k-means cluster analysis

```{r}
fit.km <- kmeans(pharmaceutical_data.norm, 5, nstart=10)
fit.km$size
fit.km$centers
fit.km$withinss
```

# calcualte cluster centroidsfit.km\$centers

```{r}
fit.km$centers


fit.km$withinss
```

#check if any cluster is a striking outlier

```{r}
dist(fit.km$centers)
```

#Profile plot of Centroids

# plot an empty scatter plot

```{r}
plot(c(0), xaxt = 'n', ylab = "", type = "l", ylim = c(min(fit.km$centers), max(fit.km$centers)), xlim = c(0, 9))
# label x-axes
axis(1, at = c(1:9), labels = names(pharmaceutical_data))
# plot centroids
for(i in c(1:5))lines(fit.km$centers[i,], lty = i, lwd = 2,
                      col = ifelse(i %in% c(1),"black",
                                (ifelse(i %in% c(2),"blue",
                                      (ifelse(i %in% c(3),"green",
                                            (ifelse(i %in% c(4),"red","dark grey"))))))))
                                          
text(x = 0.5, y = fit.km$centers[, 1], labels = paste("Cluster", c(1:5))) 
```

Cluster 5 can be called as \'Big doing great\' as it has the highest market cap with high Asset turnover, low Beta(risk) and a profit margin greater than all other clusters.

Cluster 3 can be called \'Recovering fast\' as it has a low market cap, high Beta(risk), lowest asset
turnover, good profit and highest revenue growth.

Cluster 4 can be called as \'High risk no recovery\' as it has the highest Beta, low market cap, low
ROE, low ROA, least revenue growth and least net profit margin.

Cluster 2 can be called as \'Stable going good\' as it has the lease Beta(risk), high ROE, high ROA,
good revenue growth and high Net Profit Margin.

Cluster 1 can be called as \'Stable best buy\' as it has the least Beta just like Cluster 1 , has an
average asset turnover, average revenue growth and has the highest PE Ratio among all the
clusters which is the factor used to select which stocks to buy. PE Ratio is the ratio of current
Stock market price to the earning per share.
